2.5: OLS: Precision and Diagnostics

### 2019-10-04 content/class/11-class.Rmd
Fix slides 11, add practice 11 answers, edit class 12 page

*Tuesday, October 2, 2019*

@@

>>>>

*Tuesday, October 1, 2019*

@@

# Appendix: Robust Standard Errors in R

This, since I started using `huxtable` instead of another package (`stargazer`) to make regression tables, I have gone all in on `estimatr`'s `lm_robust()` option to calculate robust standard errors. Before this, there were some other methods that I had to resort to. You can read about that in [this blog post](http://ryansafner.com/post/replicating-statas-robust-option-for-ols-standard-errors-in-r).

### 2019-10-04 content/class/12-class.Rmd
Add problem set 3 mention in class 12
Fix slides 11, add practice 11 answers, edit class 12 page

*Thursday, October 3, 2019*

@@

We begin with some more time for you to work on the [R Practice]() from last class. I will post the answer key after class.

This class and the next class are about inferential statistics: using *statistics* calculated from a *sample* of data to **infer** the true (and unmeasurable) *parameters* that describe a *population.* In doing so, we can run *hypothesis tests* on our sample to determine a *point* estimate of a parameter, or construct a *confidence interval* from our sample to cast a range for the true parameter.

@@

Today you will be finishing up [R practice problems](/r/11-r-practice.html) from last class. Check back later for solutions.

@@

# Assignments: Problem Set 2 Answers

[Problem Set 2](/assignment/01-problem-set.html) answers are posted.

>>>>

*Thursday, October 3, 2019* and *Tuesday, October 8, 2019*

@@

We begin with some more time for you to work on the [R Practice](/r/11-r-practice) from last class. The answer key is now posted.

This class will be spread over 2 days, and is about inferential statistics: using *statistics* calculated from a *sample* of data to **infer** the true (and unmeasurable) *parameters* that describe a *population.* In doing so, we can run *hypothesis tests* on our sample to determine a *point* estimate of a parameter, or construct a *confidence interval* from our sample to cast a range for the true parameter.

@@

You will be finishing up [R practice problems](/r/11-r-practice.html) from last class. Answers are posted.

@@

# Assignments: Problem Set 3 Due Thursday October 10

Problem Set 3 will be posted later and is due Thursday September 19.

### 2019-10-04 content/r/11-r-practice.Rmd
Add link to data in practice 11
Fix slides 11, add practice 11 answers, edit class 12 page

---
title: "2.4 OLS: Precision and Diagnostics - R Practice (Answers)"
editor_options: 
  chunk_output_type: console
---

@@

Our [data](https://github.com/fivethirtyeight/data/tree/master/congress-trump-score) comes from fivethirtyeight's [Trump Congress tracker](https://projects.fivethirtyeight.com/congress-trump-score/).

- [View Project on R Studio Cloud](https://rstudio.cloud/project/575252)

## Question 1
Download and read in (`read_csv`) the data.

@@

### Part C
Look in your original data to match this outlier with an observation. Hint: use the `slice()` command, as the outlier test gave you an observation number!

@@

This data is still a bit messy. We'd also probably like to plot our scatterplots with colors for Republican and Democratic party. Or sort by the House and the Senate.

>>>>

---
title: "2.4 OLS: Precision and Diagnostics - R Practice"
editor_options: 
  chunk_output_type: console
---

@@

# Set Up
To minimize confusion, I suggest creating a new `R Project` (e.g. `regression_practice`) and storing any data in that folder on your computer.

Alternatively, I have made a project in R Studio Cloud that you can use (and not worry about trading room computer limitations), with the data already inside (you will still need to assign it to an object).

- [<i class="fas fa-cloud"></i> View Project on R Studio Cloud](https://rstudio.cloud/project/575252)
- [<i class="fab fa-markdown"></i> `Download Markdown`](/r/11-r-practice-rmd.Rmd)

- [<i class="fas fa-external-link-square-alt"></i> `Answer Key (html)`](/r/11-r-practice-answers.html)
- [<i class="fab fa-markdown"></i> `Answer Key (markdown)`](/r/11-r-practice-answers.Rmd)

## Question 1
Our [data](https://github.com/fivethirtyeight/data/tree/master/congress-trump-score) comes from fivethirtyeight's [Trump Congress tracker](https://projects.fivethirtyeight.com/congress-trump-score/). Download and read in (`read_csv`) the data.

- [<i class="fas fa-table"></i> `congress-trump-score.csv`](/data/congress-trump-score.csv)

@@

### Part C
Look in your original data to match this outlier with an observation. Hint: use the `slice()` command, as the outlier test gave you an observation (row) number!

@@

This data is still a bit messy. Let's check in on your `tidyverse` skills again! For example, we'd probably like to plot our scatterplots with colors for Republican and Democratic party. Or plot by the House and the Senate.

### Part A
First, the variable `congress` (session of Congress) seems a bit off. Get a `count()` of `congress`.

### Part B
Let's get rid of the `0` values for `congress` (someone made a mistake coding this, probably). Also, while we're at it, let's take `agree_pct` and `mutate` a variable that is a proper percentage (i.e. `*100`).

### Part C
The variable `party` is also quite a mess. `count()` by `party` to see. Then let's `mutate` a variable to make a better measure of political party - just `"Republican"`, `"Democrat"`, and `"Independent"`. Try doing this with the `case_when()` command (as your `mutate` formula).^[The syntax for `case_when()` is to have a series of `condition ~ "Outcome"`, separated by commas. For example, one condition is to assign both `"Democrat"` and `"D"` to `"Democrat"`, as in `party %in% c("Democrat", "D") ~ "Democrat"`. You could also do this with a few `ifelse()` commands, but that's a bit more awkward.] When you're done `count()` by your new party variable to make sure it worked.

### Part D
Now plot a scatterplot (same as Question 3) and set `color` to your party variable. Notice `R` uses its own default colors, which don't match to the actual colors these political parties use! Make a vector where you define the party colors as follows: `party_colors <- c("Democrat" = "blue", "Republican" = "red", "Independent" = "gray")`. Then, run your plot again, adding the following line to customize the colors `+scale_colour_manual("Parties", values = party_colors)`.^[`"Parties"` is the title that will show up on the legend, feel free to edit it, or remove the legend with another layer `+guides(color = F)`.]

### Part E
Now facet your scatterplot by `chamber`.

### 2019-10-04 static/slides/11-slides.Rmd
Fixed slides 11
Fix slides 11, add practice 11 answers, edit class 12 page

.pull-right[
.font80[
.regtable[
```{r}
library(huxtable)
huxreg("Test Score" = school_reg,
       coefs = c("Intercept" = "(Intercept)",
                 "STR" = "str"),
       statistics = c("N" = "N",
                      "R-Squared" = "r.squared",
                      "SER" = "sigma"),
       number_format = 2)
```
]
]
]

@@

```{r, echo=T, eval=F}
statistics = c("N" = "N",
               "R-Squared" = "r.squared",
               "SER" = "sigma")
```

@@

```{r, echo=T, eval=F}
huxreg("Test Score" = school_reg,
       coefs = c("Intercept" = "(Intercept)",
                 "STR" = "str"),
       statistics = c("N" = "N",
                      "R-Squared" = "r.squared",
                      "SER" = "sigma"),
       number_format = 2)
```

@@

.pull-left[
.font80[
.regtable[
```{r}
huxreg("Test Score" = school_reg,
       coefs = c("Intercept" = "(Intercept)",
                 "STR" = "str"),
       statistics = c("N" = "N",
                      "R-Squared" = "r.squared",
                      "SER" = "sigma"),
       number_format = 2)
```
]
]
]

@@

4. There is no correlation between $X$ and the error term: 
$$cor(X, u)=0 \text{ or } E[u|X]=0$$

@@

```{r, echo=T}
aug_reg %>%
  summarize(e_u = mean(.resid),
            sd_u = sd(.resid))
```

@@

ggplot(data = aug_het_reg)+
  aes(x = .resid)+
  geom_density(aes(fill=range), alpha=0.5)+
  geom_label(data=aug_het_reg_sigmas,
             aes(x=0,y=0.3,
                 label=sigmas,
                 color=range),size=5)+
                 #label=parse(text=paste('~sigma==', sigmas))))+
  facet_grid(~range)+
  guides(fill = F,
         color = F)+
  labs(x = "Residuals by Range of X",
       y = "Density")+
  theme_bw(base_family = "Fira Sans Condensed",
           base_size=20)

@@

```{r}
library(wooldridge)
wage_reg<-lm(wage~educ, data=wage1)
huxreg("Wage" = wage_reg,
       coefs = c("Intercept" = "(Intercept)",
                 "Years of Schooling" = "educ"),
       statistics = c("N" = "N",
                      "R-Squared" = "r.squared",
                      "SER" = "sigma"),
       number_format = 2)
```
]

@@

```{r, fig.height=3.5, fig.retina=3}
aug_wage_reg<-aug_wage_reg %>%
  mutate(range = case_when(educ>=0 & educ<4 ~ "0 to 4",
                           educ>=4 & educ<8 ~ "4 to 8",
                           educ>=8 & educ<12 ~ "8 to 12",
                           educ>=12 & educ<16 ~ "12 to 16",
                           educ>=16 & educ<20 ~ "16 to 18"),
         range = factor(range, levels = c("0 to 4", "4 to 8", "8 to 12", "12 to 16", "16 to 18"))) # needs to be a factor to be in order in the plot

@@

```{r}
#install.packages("estimatr")
library(estimatr)

@@

.pull-left[
```{r, echo=T, eval=F}
library(huxtable)
huxreg("Normal" = school_reg,
       "Robust" = school_reg_robust,
       coefs = c("Intercept" = "(Intercept)",
                 "STR" = "str"),
       statistics = c("N" = "N",
                      "R-Squared" = "r.squared",
                      "SER" = "sigma"),
       number_format = 2)

@@

.pull-right[
.font80[
.regtable[
```{r}
huxreg("Normal" = school_reg,
       "Robust" = school_reg_robust,
       coefs = c("Intercept" = "(Intercept)",
                 "STR" = "str"),
       statistics = c("N" = "N",
                      "R-Squared" = "r.squared",
                      "SER" = "sigma"),
       number_format = 2)
```
]
]
]

@@

huxreg("No Outliers" = school_reg,
       "Outliers" = school_outlier_reg,
       coefs = c("Intercept" = "(Intercept)",
                 "STR" = "str"),
       statistics = c("N" = "N",
                      "R-Squared" = "r.squared",
                      "SER" = "sigma"),
       number_format = 2)
```
]
]

>>>>

.pull-right[
.font80[
.regtable[
```{r}
library(huxtable)
huxreg("Test Score" = school_reg,
       coefs = c("Intercept" = "(Intercept)",
                 "STR" = "str"),
       statistics = c("N" = "nobs",
                      "R-Squared" = "r.squared",
                      "SER" = "sigma"),
       number_format = 2)
```
]
]
]

@@

```{r, echo=T, eval=F}
statistics = c("N" = "nobs",
               "R-Squared" = "r.squared",
               "SER" = "sigma")
```

@@

```{r, echo=T, eval=F}
huxreg("Test Score" = school_reg,
       coefs = c("Intercept" = "(Intercept)",
                 "STR" = "str"),
       statistics = c("N" = "nobs",
                      "R-Squared" = "r.squared",
                      "SER" = "sigma"),
       number_format = 2)
```

@@

.pull-left[
.font80[
.regtable[
```{r}
huxreg("Test Score" = school_reg,
       coefs = c("Intercept" = "(Intercept)",
                 "STR" = "str"),
       statistics = c("N" = "nobs",
                      "R-Squared" = "r.squared",
                      "SER" = "sigma"),
       number_format = 2)
```
]
]
]

@@

4. No correlation between $X$ and the error term: 
$$cor(X, u)=0 \text{ or } E[u|X]=0$$

@@

```{r, echo=T}
aug_reg %>%
  summarize(E_u = mean(.resid),
            sd_u = sd(.resid))
```

@@

ggplot(data = aug_het_reg)+
  aes(x = .resid)+
  geom_density(aes(fill=range), alpha=0.5)+
  geom_label(data=aug_het_reg_sigmas,
             aes(x=0,y=0.3,
                 label=sigmas,
                 color=range),size=5)+
                 #label=parse(text=paste('~sigma==', sigmas))))+
  facet_grid(~range)+
  guides(fill = F,
         color = F)+
  labs(x = "Residuals by Range of X",
       y = "Density")+
  theme_bw(base_family = "Fira Sans Condensed",
           base_size=16)

@@

```{r}
library(wooldridge)
wage_reg<-lm(wage~educ, data=wage1)
huxreg("Wage" = wage_reg,
       coefs = c("Intercept" = "(Intercept)",
                 "Years of Schooling" = "educ"),
       statistics = c("N" = "nobs",
                      "R-Squared" = "r.squared",
                      "SER" = "sigma"),
       number_format = 2)
```
]

@@

aug_wage_reg<-aug_wage_reg %>%
  mutate(range = case_when(educ>=0 & educ<4 ~ "0 to 4",
                           educ>=4 & educ<8 ~ "4 to 8",
                           educ>=8 & educ<12 ~ "8 to 12",
                           educ>=12 & educ<16 ~ "12 to 16",
                           educ>=16 & educ<20 ~ "16 to 18"),
         range = factor(range, levels = c("0 to 4", "4 to 8", "8 to 12", "12 to 16", "16 to 18"))) # needs to be a factor to be in order in the plot

@@

```{r, echo = T}
#install.packages("estimatr")
library(estimatr)

@@

.pull-left[
```{r, echo=T, eval=F}
library(huxtable)
huxreg("Normal" = school_reg,
       "Robust" = school_reg_robust,
       coefs = c("Intercept" = "(Intercept)",
                 "STR" = "str"),
       statistics = c("N" = "nobs",
                      "R-Squared" = "r.squared",
                      "SER" = "sigma"),
       number_format = 2)

@@

.pull-right[
.font80[
.regtable[
```{r}
huxreg("Normal" = school_reg,
       "Robust" = school_reg_robust,
       coefs = c("Intercept" = "(Intercept)",
                 "STR" = "str"),
       statistics = c("N" = "nobs",
                      "R-Squared" = "r.squared",
                      "SER" = "sigma"),
       number_format = 2)
```
]
]
]

@@

huxreg("No Outliers" = school_reg,
       "Outliers" = school_outlier_reg,
       coefs = c("Intercept" = "(Intercept)",
                 "STR" = "str"),
       statistics = c("N" = "nobs",
                      "R-Squared" = "r.squared",
                      "SER" = "sigma"),
       number_format = 2)
```
]
]
